---
title: "Keyword Assited Topic Models - Demonstration"
author: "PLSC 21510/31510"
date: "Fall 2022"
output: html_document
---

This unit gives a brief overview of the `keyATM` (keyword-assited topic models) package. Please read [the vignette] (https://keyatm.github.io/) for more detail.

## 1. Install

```{r}
# install first
# install.packages("keyATM")
require(keyATM)
require(quanteda)
require(tidyverse)
```

## 3. Preparing texts

Since `keyATM` does not provide preprocessing functions, we recommend users to preprocess texts with `quanteda.` 

```{r}
data(data_corpus_inaugural)
data_tokens <- tokens(data_corpus_inaugural,
                      remove_numbers = TRUE, 
                      remove_punct = TRUE, 
                      remove_symbols = TRUE,
                      remove_separators = TRUE,
                      remove_url = TRUE) %>%
  tokens_tolower() %>%
  tokens_remove(c(stopwords("english"), 
                  "may", "shall", "can",
                  "must", "upon", "with", "without")) %>%
  tokens_select(min_nchar = 3)

# create DFM
data_dfm <- dfm(data_tokens) %>%
  dfm_trim(min_termfreq = 5, min_docfreq = 2)

# Finally, `keyATM_read` function reads your data for keyATM.
keyATM_docs <- keyATM_read(texts = data_dfm)
summary(keyATM_docs)
```

## 3. Preparing keywords 

Suppose we are interested in five topics: Government, Congress, Peace, Constitution, and Foreign affairs. We use our substantive knowledge to find keywords for these topics. 

Keywords should be stored in a list object. Each keyword-topic can have a different number of keywords.

```{r}
keywords <- list(Government     = c("laws", "law", "executive"),
                 Congress       = c("congress", "party"),
                 Peace          = c("peace", "world", "freedom"),
                 Constitution   = c("constitution", "rights"),
                 ForeignAffairs = c("foreign", "war"))
```


Keywords should appear reasonable times (> 0.1% of the corpus) in the documents. The `visualize_keywords()` function plots the frequency of keywords by topic. Get the actual values with the `values_fig()` 

```{r}
function.key_viz <- visualize_keywords(docs = keyATM_docs, keywords = keywords)
key_viz
values_fig(key_viz) 
```

## 3. Fitting the model

The next step is to fit a model with the keyATM() function. keyATM has three models:

1. keyATM Base: This is an extension of the most famous topic model, Latent Dirichlet Allocation. If you do not have covariates, this model is your first option.

2. keyATM Covariates: If you have covariates, please use this model. This model uses document-level meta data (document-level covariates) to model topic prevalence (the prior of document-topic distribution).

3. keyATM Dynamic: If you want to explicitly consider time structure, please use this model.

### Base model

Let's use #1 for now.

We pass the output of the `keyATM_read` function and keywords to the `keyATM` function.

Additionally, we need to specify the number of topics without keywords (the no_keyword_topics argument) and model. Since this example does not use covariates or time stamps, base is the appropriate model.

To guarantee the replicability, we recommend to set the random seed in the option argument (see here for other options). The default number of iterations is 1500.

```{r}
out <- keyATM(docs              = keyATM_docs,    # text input
              no_keyword_topics = 5,              # number of topics without keywords
              keywords          = keywords,       # keywords
              model             = "base",         # select the model
              options           = list(seed = 250))
```

### Saving the model
Once you fit the model, you can save the model with the save() function for replication. We strongly recommend to save the fitted model.

```{r}
save(out, file = "SAVENAME.rds")

# To load the model, you can use readRDS() function.
out <- readRDS(file = "SAVENAME.rds")
```

### Interpreting results

The `top_words()` function returns a table of top words for each of estimated topics. Keywords assigned to a keyword topic are suffixed with a check mark. Keywords from another keyword topic are labeled with the topic id of that category.

In the table below, "law", "laws", and "executive" are keywords of the Government topic, while "peace" appears in top words of the Other_3 topic, it is a keyword of the peace topic.

```{r}
top_words(out)
```

To explore documents that are highly associated with each topic, the `top_docs()` function returns a table of document indexes in which a topic has high proportion.

The table below indicates, for example, that the ninth document in the corpus has the highest proportion of the Government topic among all other documents.

```{r}
top_docs(out)
```

Researchers may want to obtain the entire document-topic distribution and topic-word distribution. The output of the `keyATM()` function contains both quantities.

```{r eval=FALSE}
out$theta  # Document-topic distribution
out$phi    # Topic-word distribution
```

## Keyword Dynamic

Researchers are often interested in investigating how the prevalence of topics change over time. Researchers should construct a numeric or an integer vector of time stamps after preprocessing texts.

Below,tThe time index (`Period`) should start from 1 and increment by 1 (ascending order). To properly process date, please consider using the `lubridate` package.

```{r eval=FALSE}
# Get meta information (time stamp)
vars <- docvars(data_corpus_inaugural)
head(vars)

# Divide by a decade
# Timestamp should start with 1 (the variable "Period")
vars_period <- vars %>%
  as_tibble() %>%
  mutate(Period = (vars$Year - 1780) %/% 10 + 1) 
head(vars_period)
```

We pass the time index to the `keyATM()` function, specifying the number of hidden states with the `num_states` in the `model_settings` argument.

```{r}
out <- keyATM(docs              = keyATM_docs,                         
              no_keyword_topics = 3,                                   
              keywords          = keywords,                      
              model             = "dynamic",                           
              model_settings    = list(time_index = vars_period$Period,
                                       num_states = 5),                
              options           = list(seed = 250))
```

We can use the` top_words()`, `top_docs()` functions as in the base `keyATM.`

```{r}
top_words(out)
top_docs(out)
```

Finally, we can plot the time trend of topics with the `plot_timetrend()` function. With the time_index_label argument, you can label each time index. Note that store_theta option in the `keyATM()` function should be TRUE to show 90% credible intervals.

```{r}
out <- keyATM(docs              = keyATM_docs,                         
              no_keyword_topics = 3,                                   
              keywords          = keywords,                      
              model             = "dynamic",                           
              model_settings    = list(time_index = vars_period$Period,
                                       num_states = 5),                
              options           = list(seed = 250, store_theta = TRUE))

fig_timetrend <- plot_timetrend(out, show_topic = c(1:5), time_index_label = vars$Year, xlab = "Year")
fig_timetrend
```