---
title: "Strings and Regular Expressions"
author: "PLSC 21510/31510"
date: "Fall 2022"
output:
  xaringan::moon_reader:
    css: ["default",  "rladies", "rladies-fonts", "libs/rt.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      highlightSpans: true

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = T,
  cache = F,
  out.width = "70%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold"
)
```

```{r include=FALSE} 
# include any code here you don't want to show up in the document,
# e.g. package and dataset loading
library(tidyverse)
library(kableExtra)
```


### Strings and Regular Expressions

This unit focuses on character (or "string") data. We'll explore:

1. .accent[string basics], like concatenating and subsetting.
2. .accent[regular expressions], a powerful cross-language tool for working with string data.
3. .accent[common tools], that take regex and apply them to real problems.

We will use the **`stringr`** package for string manipulation. `stringr` is part of `tidyverse` but we need to load it explicitly.

```{r message = FALSE}
library(tidyverse)
library(stringr)
```

---
class: center, middle, inverse, title-slide

# String Basics


---
### Creating Strings

Create strings with either single quotes or double quotes. 

```{r}
string1 <- "This is a string"
```

--
Use `"` unless you want to create a string that contains multiple `"`.

```{r}
string2 <- 'If I want to include a "quote" inside a string, I use single quotes'
```

--
Multiple strings are stored in a character vector.

```{r}
string_vector <- c("one", "two", "three")
```

---
### Escape and Special Characters 

Single and double quotes are known as .accent[metacharacters], meaning that they have special meaning to the R language. 

To include a .accent[literal] single or double quote in a string you can use `\` to "escape" it:

```{r}
double_quote <- "\"" # or '"'
single_quote <- '\'' # or "'"

# To include a literal backslash, double it up
backslash <- "\\"
```

---
### Printed Representation vs. Raw

The .accent[printed representation] of a string is not the same as the string itself, because the printed representation shows the escapes. 

To see the raw contents of the string, use `writeLines()`:

```{r}
x <- c("\"", "\\")
x

writeLines(x)
```

---
### Other special characters

#### Most common:
* `"\n"`: newline, 
* `"\t"`, tab, 
* See the complete list by requesting help on `"`: `?'"'`.

#### Unicode:
* Strings like `"\u00b5"`: non-English characters that works on all platforms.

```{r}
x <- "\u00b5"
x
```

---
### Measure string length with `str_length()`

All `stringr` functions start with `str_` (helpful to trigger autocomplete). 

For example, `str_length()` tells you the number of characters in a string:

```{r}
str_length(c("a", "R for data science", NA))
```

---
### Combine strings with `str_c()`

To combine two or more strings, use `str_c()`:

```{r}
str_c("x", "y")

str_c("x", "y", "z")
```

--
Use the `sep` argument to control how they are separated:

```{r}
str_c("x", "y", sep = ", ")
```

--
`str_c()` is vectorised.

```{r}
x <- c("a", "b", "c")
str_c("prefix-", x)
```

---
### Subset strings with `str_sub()`

`str_sub()` takes `start` and `end` arguments, which give the (inclusive) position of the substring.

```{r}
x <- c("Rochelle is the GOAT")
str_sub(x, 1, 8)

# Negative numbers count backwards from the end
str_sub(x, -8, -1)
```

--
You can also use the assignment form of `str_sub()` to modify strings:

```{r}
x <- c("Rochelle is the GOAT")
str_sub(x, 1, 1) <- str_to_lower(str_sub(x, 1, 1))
x
```

---
### Locales

`str_to_lower()`, `str_to_upper()`, and `str_to_title()` changes case. 

You can pick which set of rules to use by specifying a **`locale`**.

The locale is specified as a ISO 639 language code, which is a two or three letter abbreviation. 

```{r}
# Turkish has two i's: 
# with and without a dot, 
# and has a different rules 
# for capitalising them:
str_to_upper(c("i", "ı"))

str_to_upper(c("i", "ı"), 
             locale = "tr")
```


---

class: center, middle, inverse, title-slide

# Regular Expressions

---
class: center, middle

#### Find a friend and complete this tutorial together: https://regexone.com/

---
### Regular expressions

.accent[Regular expressions] (.accent[regex] or .accent[regexps]) are a very terse language that allow you to describe patterns in strings. 

To learn regular expressions, we'll use `str_view()` and `str_view_all()`. These functions take a character vector and a regular expression, and show you how they match. 

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
```

---
### Basic matches

.pull-left[
The simplest patterns match exact strings:

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
```
]

.pull-right[
**`.`** matches any character (except a newline):

```{r}
x <- c("apple", "banana", "pear")
str_view(x, ".a.")
```
]

---
### Escape Characters

If **`.`** matches any character, how do you match the literal character "**`.`**"? 

Regexps use the backslash, `\`, to escape metacharacters. So to match an **`.`**, you need the regexp `\.` 

But this creates a problem! We use *strings* to represent regular expressions, and `\` is also used as an escape symbol in strings. 

So to create the regular expression `\.` we need the string `"\\."`. 

```{r}
# To create the regular expression, we need \\
dot <- "\\."

# But the expression itself only contains one:
writeLines(dot)
```

---
### Anchors

> "if you begin with power (`^`), you end up with money (`$`).""

.pull-left[
Use `^` to match the start of the string.

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
```
]


.pull-right[
Use `$` to match the end of the string.

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "a$")
```
]

---
### Anchors

To force a regular expression to only match a complete string, anchor it with both `^` and `$`:

.pull-left[
```{r}
x <- c("apple pie", "apple", 
       "apple cake")
str_view(x, "apple")
```
]

.pull-right[
```{r}
x <- c("apple pie", "apple", 
       "apple cake")
str_view(x, "^apple$")
```
]

---
### Character classes

There are a number of special patterns that match more than one character. 

* `\d`: matches any digit.
* `\s`: matches any whitespace (e.g. space, tab, newline).
* `[abc]`: matches a, b, or c.
* `[^abc]`: matches anything except a, b, or c..
* `[A-Z]`: matches capital letters A through Z.

.pull-left[

```{r}
x <- c("a1c", "a.c", "a*c", "a c")
str_view(x, "\\d")
```
]

.pull-right[

```{r}
x <- c("a1c", "a.c", "a*c", "a c")
str_view(x, "a[.]c")
```
]

---
### Alternation

Use _alternation_ to pick between one or more alternative patterns. 

For example, `abc|deaf` will match either '"abc"', or `"deaf"`. 

Like with mathematical expressions, use parentheses to make it clear what you want.

```{r}
x <- c("grey", "gray")
str_view(x, "gr(e|a)y")
```

---
### Repetition

Control how many times a pattern matches:

* `?`: 0 or 1
* `+`: 1 or more
* `*`: 0 or more

.pull-left[
```{r}
x <- "MDCCCLXXXVIII"
str_view(x, "CC?")
```
]

.pull-right[
```{r}
x <- "MDCCCLXXXVIII"
str_view(x, 'C[LX]+')
```
]

---
class: center, middle, inverse, title-slide

# Applying Regex

---
## Applying regex

Let's apply regex using `stringr` functions that let you

* Detect matches in a string with `str_detect()`.
* Count the number of matches with `str_count()`
* Extract matches with `str_extract()` and `str_extract_all()`
* Replace matches with `str_replace()` and `str_replace_all()`
* Split a string based on a match with `str_split()`

---
### Detect matches: `str_detect()`

To determine if a character vector matches a pattern, use `str_detect()`. It returns a logical vector the same length as the input:

```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")
```

Remember that, when you use a logical vector in a numeric context, `FALSE` becomes 0 and `TRUE` becomes 1.
```{r}
# Get common words
words <- stringr::words
# How many common words start with t?
sum(str_detect(words, "^t"))
# What proportion end with a vowel?
mean(str_detect(words, "[aeiou]$"))
```

---
### Subset matches: `str_subset()`

To select the elements that match a pattern, use with `filter()`

```{r}
df <- data.frame(
  i = seq_along(words),
  word = words
)
df %>% 
  filter(str_detect(word, "x$"))
```

---
### Count matches: `str_count()`

* To count how many matches there are in a string, use `str_count()`

```{r}
df1 <- df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )

head(df1)
```

---
### Extract matches:  `str_extract()`

To extract the actual text of a match, use `str_extract()`.

```{r}
head(sentences)
```

.accent[How do we find all sentences that contain a color?]

---
### Extract matches:  `str_extract()`

```{r}
# create a 'color' regex
colors <- c("red", "orange", "yellow", "green", "blue", "purple")
color_match <- str_c(colors, collapse = "|")
color_match
```

--
```{r}
# find sentences with colors
has_color <- str_subset(sentences, color_match)
head(has_color, 3)
```

--
```{r}
# extract the color
matches <- str_extract(has_color, color_match)
head(matches)
```

---
### Extract matches: : `str_extract()`

Note that `str_extract()` only extracts the first match. 

This is a common pattern for stringr functions, because working with a single match allows you to use much simpler data structures. 

To get all matches, use `str_extract_all()`. It returns a list:

```{r}
all_colors <- str_extract_all(has_color, color_match)
all_colors[18:20]
```

---
### Extract matches: `str_extract()`

If you use `simplify = TRUE`, `str_extract_all()` will return a matrix with short matches expanded to the same length as the longest:

```{r}
color_mat <- str_extract_all(has_color, color_match, simplify = TRUE)
color_mat[18:22,]
```

---
### Replacing matches: `str_replace()`

`str_replace()` and `str_replace_all()` allow you to replace matches with new strings. 

```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-") # replace the first instance of a match
str_replace_all(x, "[aeiou]", "-") # replace all instances of a match
```

--
With `str_replace_all()` you can perform multiple replacements by supplying a named vector:

```{r}
x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```

---
### Splitting: `str_split()`

Use `str_split()` to split a string up into pieces. 

```{r}
sentences %>%
  str_split(" ") %>%
  head(3)
```

---
### Splitting: `str_split()`

Like the other `stringr` functions that return a list, you can use `simplify = TRUE` to return a matrix:

```{r}
sentences %>%
  str_split(" ", simplify = TRUE) %>%
  head(3)
```

---
### Splitting: `str_split()`

Instead of splitting up strings by patterns, you can also split up by character, line, sentence and word `boundary()`s:

```{r}
x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))
str_split(x, boundary("word"))[[1]]
```
